### 2024-03-12 cover_letter/cover_letter.tex
Update cover letter

\documentclass{letter}

\signature{Daiki Tagami}

\address{Big Data Institute\\University of Oxford\\UK}
\begin{document}

\begin{letter}{Bioinformatics}

\opening{Dear Editors,}

I am writing on behalf of my coauthors to submit our manuscript entitled
\emph{tstrait: a quantitative trait simulator for ancestral recombination graphs}
for consideration for publication in Bioinformatics
as an ``Application Note'' research article.

Following recent breakthroughs in simulation and inference,
Ancestral Recombination Graphs (ARGs) are the subject of
intense research interest (see Lewanski et al 2024, PLoS Genetics,
and Brandt et al 2024, GBE, for recent perspectives). One of the key application
areas is statistical genetics, and in particular developing
improved methodology for Genome Wide Association Studies (GWAS).
However, there is currently no available software to efficiently
and conveniently simulate quantitative traits from an ARG,
significantly hampering progress. In this manuscript we describe
\texttt{tstrait}, a new Python package that can efficiently
simulate quantitative traits on ARGs under a wide range of
models. We believe that this tool will be of significant interest
to the readers of Bioinformatics, and will enable further
ground-breaking research in this area.

\closing{Sincerely,}

\end{letter}
\end{document}