### 2024-05-13 cover_letter/cover_letter_resubmission.tex

\documentclass{letter}

\signature{Daiki Tagami}

\address{Big Data Institute\\University of Oxford\\UK}
\begin{document}

\begin{letter}{Bioinformatics}

\opening{Dear Editors,}

I am writing on behalf of my coauthors to resubmit our manuscript entitled
\emph{tstrait: a quantitative trait simulator for ancestral recombination graphs}
for consideration for publication in Bioinformatics
as an ``Application Note'' research article.

We thank the reviewers for their thoughtful and thorough reviews, and we have addressed those points in the revised manuscript. We have attached
a detailed point-by-point response in the \texttt{response-to-reviewers.pdf}
file.

We have also attached a file showing the changes made to the manuscript
in response to reviewers in \texttt{review-diff.pdf}.

Finally, we attach \texttt{supplementary-material.pdf}, which includes some clarifications on our 
statistical testing methodology.

\closing{Sincerely,}

\end{letter}
\end{document}